---
title: "Class Level Fidelity"
author: "Kirk Vanacore"
date: "9/15/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, include=FALSE}
require(psych)
require(ggplot2) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)
library(lme4)
library(stringr)
ifnull <- function(x = NULL, alt = NULL){
  if(is.null(x)){
    x <- alt
  }
  x
}
```

# Fidelity Metric

> Science  cannot  study  what  it  cannot  measure  accurately  and  cannot  measure  what  it  does  not  define (Durlak & DuPre, 2008).

* Fidelity is using the program as it was intended
* In this study parctitioners were expected to to assigning the online assignments and create time for students to utilize the program
* Our working definition of minumim fedility is whether this expection was met
* Because the action of assigning metrics was not measured directly, we utilize whether at least 25% students within a class started each assignment as our threshould for minimum fedility
 


```{r, include=F, echo=F}
assess <-read.csv("Assessment_merged_2021_07_16_state_assessment_N=4321 - Sheet1.csv", na.strings = c("", "N/A"))

# Build Analytic Sample 
  # schools who did not implement: S03 (for dropping before completing the pre-test) and S07 
  # 
table(assess$initial_school_id)
table(assess$final_school_id)

length(unique(assess$student_id)) # 4193
table(is.na(assess$rdm_condition)) # 4092 original assigned to random treatment group
table((assess$rdm_condition)) # 120 in resource 
levels((assess$rdm_condition)) # 120 in resource 




assess_clean <- assess %>%
  filter(
    # remove students who are not in original random assignment
    is.na(assess$rdm_condition) == F   
    # remove students who are in resource
    & assess$rdm_condition != "Dragon-Resource" & assess$rdm_condition != "FH2T-Resource"   
    # S03 drop schools 
    & final_school_id != "S03" & initial_school_id != "S03"
    & final_school_id != "S07" & initial_school_id != "S07"

  )
assess <- assess_clean
rm(assess_clean)
```

## Class Level Fidelity Metrics
```{r}
# create data file with one row per class for fidelity metrics
class <- assess %>%
  ungroup() %>%
  group_by(final_teacher_class, final_school_id) %>%
  dplyr::summarise(num_students = n(),
            num_started_assignment_2= sum(ifnull(started_assignment_2),0),
            num_started_assignment_3= sum(ifnull(started_assignment_3),0),
            num_started_assignment_4= sum(ifnull(started_assignment_4),0),
            num_started_assignment_5= sum(ifnull(started_assignment_5),0),
            num_started_assignment_7= sum(ifnull(started_assignment_7),0),
            num_started_assignment_8= sum(ifnull(started_assignment_8),0),
            num_started_assignment_9= sum(ifnull(started_assignment_9),0),
            num_started_assignment_10= sum(ifnull(started_assignment_10),0),
            num_started_assignment_11= sum(ifnull(started_assignment_11),0),
            pre_started_assignment_2= mean(started_assignment_2),
            pre_started_assignment_3= mean(started_assignment_3),
            pre_started_assignment_4= mean(started_assignment_4),
            pre_started_assignment_5= mean(started_assignment_5),
            pre_started_assignment_7= mean(started_assignment_7),
            pre_started_assignment_8= mean(started_assignment_8),
            pre_started_assignment_9= mean(started_assignment_9),
            pre_started_assignment_10= mean(started_assignment_10),
            pre_started_assignment_11= mean(started_assignment_11)
            ) %>%
  mutate(
    pre_assignments_started = 
      (num_started_assignment_2 +
      num_started_assignment_3 +
      num_started_assignment_4 +
      num_started_assignment_5 +
      num_started_assignment_7 +
      num_started_assignment_8 +
      num_started_assignment_9 +
      num_started_assignment_10 +
      num_started_assignment_11 )/(num_students*9)
  )

```
## Calculating Fedility Metric
```{r}
length(unique(class$final_teacher_class))
table(class$num_students)
# why are there classes with 56 students?
# there are also 69 students without a final class id

# overall percent students started lesson in class
describe(class$pre_assignments_started)
ggplot(
  data = class,
  aes(x = pre_assignments_started)) +
  geom_density(adjust = 1.5, alpha = .7,  fill = "#00BFC4") +
  theme_minimal() +
  labs(title = "Avg Session Time by Condition",
       x = "Avg Minutes per Assignment") +
  theme(plot.title = element_text(hjust = 0.5)
        ) +
  geom_vline(xintercept=.25, 
                color = "Blue", size=2)

# number who met this metric
table(class$pre_assignments_started > .25)


# fidelity for each assignment
class <- class %>%
  mutate(a2_fedility = ifelse(pre_started_assignment_2 >.25, 1, 0),
         a3_fedility = ifelse(pre_started_assignment_3 >.25, 1, 0),
         a4_fedility = ifelse(pre_started_assignment_4 >.25, 1, 0),
         a5_fedility = ifelse(pre_started_assignment_5 >.25, 1, 0),
         a7_fedility = ifelse(pre_started_assignment_7 >.25, 1, 0),
         a8_fedility = ifelse(pre_started_assignment_8 >.25, 1, 0),
         a9_fedility = ifelse(pre_started_assignment_9 >.25, 1, 0),
         a10_fedility = ifelse(pre_started_assignment_10 >.25, 1, 0),
         a11_fedility = ifelse(pre_started_assignment_11 >.25, 1, 0),
         )
table(class$a2_fedility)
table(class$a3_fedility)
table(class$a4_fedility)
table(class$a5_fedility)
table(class$a7_fedility)
table(class$a8_fedility)
table(class$a9_fedility)
table(class$a10_fedility)
table(class$a11_fedility)
      
class$overall_fideility<-ifelse(
      (class$a2_fedility) ==1&
      (class$a3_fedility) ==1&
      (class$a4_fedility) ==1&
      (class$a5_fedility) ==1&
      (class$a7_fedility) ==1& 
      (class$a8_fedility) ==1& 
      (class$a9_fedility) ==1& 
      (class$a10_fedility) ==1&
      (class$a11_fedility) ==1, 1, 0)
table(class$overall_fideility, class$pre_assignments_started > .25)
table(is.na(class$overall_fideility))
table(is.na(class$overall_fideility))

mean(class$a2_fedility, na.rm = T)
mean(class$a3_fedility, na.rm = T)
mean(class$a4_fedility, na.rm = T)
mean(class$a5_fedility, na.rm = T)
mean(class$a7_fedility, na.rm = T)
mean(class$a8_fedility, na.rm = T)
mean(class$a9_fedility, na.rm = T)
mean(class$a10_fedility, na.rm = T)
mean(class$a11_fedility, na.rm = T)

assess <- assess %>%
  left_join(
    class %>%
      select(
        final_teacher_class, 
        final_school_id,
        overall_fideility,
        pre_assignments_started
      ),
    by = c("final_school_id", "final_teacher_class")
  )

# there is almost of missing data here
table(is.na(assess$overall_fideility))
table(is.na(assess$post.percentage_math_score),is.na(assess$overall_fideility))
table(is.na(assess$pre.percentage_math_score),is.na(assess$overall_fideility))
```

## Pretest Scores by Fidelity 
```{r}
ggplot(
  assess[is.na(assess$post.percentage_math_score) == F &
           is.na(assess$overall_fideility) == F, ],
  aes(x =  as.factor(overall_fideility), 
      y = pre.percentage_math_score,
      group = overall_fideility,
      
      )
) +
  labs(title ="Pre Test Scores by Fideility",
    #   subtitle = "",
       y = "Pretest Score",
       x = "Class Level Fidelity ") +
  geom_boxplot(alpha = .7) +
  theme_minimal()

ggplot(
  assess[is.na(assess$post.percentage_math_score) == F &
           is.na(assess$overall_fideility) == F, ],
  aes(x = as.factor(overall_fideility), 
      y = pre.percentage_math_score,
  #    grpup = as.factor(virtual),
      fill = as.factor(virtual)
      
      )
) +
  labs(title ="Pre Test Scores by Fideility",
    #   subtitle = "",
       y = "Pretest Score",
       x = "Class Level Fidelity ") +
  geom_boxplot(alpha = .7) +
  theme_minimal()


```

