---
title: "Fidelity Metrics Analysis"
author: "Kirk Vanacore"
date: "9/15/2021"
output: html_document
---

The following analyses examine the cross condition fidelity metrics: number of assignments started and completed during the experiment. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, include=FALSE}
require(psych)
require(ggplot2) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)
library(lme4)
library(stringr)
```

# IES Assessment Data

```{r}
assess <-read.csv("Assessment_merged_2021_07_16_state_assessment_N=4321 - Sheet1.csv", na.strings = c("", "N/A"))
colnames(assess)

# Build Analytic Sample 
  # schools who did not implement: S03 (for dropping before completing the pre-test) and S07 
  # 
table(assess$initial_school_id)
table(assess$final_school_id)

length(unique(assess$student_id)) # 4193
table(is.na(assess$rdm_condition)) # 4092 original assigned to random treatment group
table((assess$rdm_condition)) # 120 in resource 
levels((assess$rdm_condition)) # 120 in resource 




assess_clean <- assess %>%
  filter(
    # remove students who are not in original random assignment
    is.na(assess$rdm_condition) == F   
    # remove students who are in resource
    & assess$rdm_condition != "Dragon-Resource" & assess$rdm_condition != "FH2T-Resource"   
    # S03 drop schools 
    & final_school_id != "S03" & initial_school_id != "S03"
    & final_school_id != "S07" & initial_school_id != "S07"

  )
assess <- assess_clean
rm(assess_clean)
```

## Basic Descriptives

How many students are there?
```{r}
length(unique(assess$student_id)) 
table(duplicated(assess$student_id)) 
table(is.na(assess$student_id)) 
length(unique(assess$student_number)) 
```


# Fedility Metrics Across Programs

**Variables to explore:**
* fidelity_started_sum
* fidelity_complete_sum        
* fidelity_complete_percent

**Note: fidelity_... variables are missing from the data dictionary**

### fidelity_started_sum
NEED TO FIND DEF -> number of started activities, maybe?

```{r}

### fidelity_started_sum ###
class(assess$fidelity_started_sum)
table(is.na(assess$fidelity_started_sum)) # 828 missing values
table(assess$fidelity_started_sum)
describe(assess$fidelity_started_sum)

# group differences in fidelity_started_sum

# density plot by group
ggplot(
  data = assess,
  aes(x = fidelity_started_sum,
      group = condition_assignment,
      fill = condition_assignment)
) +
  geom_density(adjust = 1.5, alpha = .4)

# between group differences
describe.by(assess$fidelity_started_sum,
            assess$condition_assignment)
summary(aov(fidelity_started_sum ~
              condition_assignment, data = assess))
TukeyHSD((
  aov(fidelity_started_sum ~
        condition_assignment, data = assess)
))

ggplot(
  assess,
  aes(x = fidelity_started_sum, y = condition_assignment, fill = condition_assignment)
) +
  geom_boxplot()

```

#### Notes:
* a lot of missing data (almost 20%)
* alot of variability
* bimodal/neg skew across conditions
* Dragonbox is sig lower than all conditions 


### fidelity_complete_sum
NEED TO FIND DEF -> number of completed activities
```{r}
### fidelity_complete_sum ###
class(assess$fidelity_complete_sum)
table(is.na(assess$fidelity_complete_sum)) # 828 missing values
table(assess$fidelity_complete_sum)
describe(assess$fidelity_complete_sum)

# group differences in fidelity_complete_sum
describeBy(assess$fidelity_complete_sum, assess$final_school_id)
describeBy(assess$fidelity_complete_sum, assess$condition_assignment)

# denisity plot by group
ggplot(
  data = assess,
  aes(x = fidelity_complete_sum,
      group = condition_assignment,
      fill = condition_assignment)
) +
  geom_density(adjust = 1.5, alpha = .4)


describe.by(assess$fidelity_complete_sum,
            assess$condition_assignment)
summary(aov(fidelity_complete_sum ~
              condition_assignment, data = assess))
TukeyHSD((
  aov(fidelity_complete_sum ~
        condition_assignment, data = assess)
))

ggplot(
  assess,
  aes(x = fidelity_complete_sum, y = condition_assignment, fill = condition_assignment)
) +
  geom_boxplot()

```

#### Notes:
* a lot of missing data (almost 20%)
  * open question: are these students who had no compliance?
* There is great variability in fidelity_complete_sum
* distributions are bimodal suggesting two groups: high and low fedelity
* Group Differences:
  * Dragonbox is significant lower from all conditions
  * see anova and tukey post hoc tests
  

### fidelity_started_sum X fidelity_complete_sum
```{r}
cor.test(
  assess$fidelity_started_sum,
  assess$fidelity_complete_sum
) 
#highly correlated



# average number of started assignments that were not completed
assess$fidelity_incomplete_sum <- assess$fidelity_started_sum - assess$fidelity_complete_sum
mean(assess$fidelity_incomplete_sum, na.rm = T)


# denisity plot by group
ggplot(
  data = assess,
  aes(x = fidelity_incomplete_sum,
      group = condition_assignment,
      fill = condition_assignment)
) +
  geom_density(adjust = 1.5, alpha = .4)

# small positive correlation between number stated and the number incomplete 
# the more the students started, the more they were likely to have incomplete
cor.test(
  assess$fidelity_started_sum,
  assess$fidelity_incomplete_sum
) 

# small negative correlation between number stated and the number incomplete 
# the more the students completed, the fewer they were likely to have incomplete
cor.test(
  assess$fidelity_complete_sum,
  assess$fidelity_incomplete_sum
) 

```
#### Notes:
* metrics are highly corrected (not surprising)
* on average, students did not complete about  
 
 
  ### fidelity_started_sum
NEED TO FIND DEF -> number of started activities, maybe?
```{r}

### fidelity_started_sum ###
class(assess$fidelity_started_sum)
table(is.na(assess$fidelity_started_sum)) # 828 missing values
table(assess$fidelity_started_sum)
describe(assess$fidelity_started_sum)

# group differences in fidelity_started_sum

# density plot by group
ggplot(
  data = assess,
  aes(x = fidelity_started_sum,
      group = condition_assignment,
      fill = condition_assignment)
) +
  geom_density(adjust = 1.5, alpha = .4)


describe.by(assess$fidelity_started_sum,
            assess$condition_assignment)
summary(aov(fidelity_started_sum ~
              condition_assignment, data = assess))
TukeyHSD((
  aov(fidelity_started_sum ~
        condition_assignment, data = assess)
))

ggplot(
  assess,
  aes(x = fidelity_started_sum, y = condition_assignment, fill = condition_assignment)
) +
  geom_boxplot()

```


# Variances of Implementation by nesting Units (School/Class)
Do schools/class differ in terms of implementation?

### Schools 

```{r}

# how many schools?
length(unique(assess$final_school_id))
table(is.na(assess$final_school_id)) # 135 have missing school ids

table((assess$final_school_id))

# school changes
table((assess$initial_school_id == assess$final_school_id)) ## 551 students moved school (this does not seem to include missing school ids)
table((assess$initial_school_id == assess$final_school_id))/length(assess$final_school_id) ## 12.8% students moved school
#will we have to deal with cross classification?

describe.by(assess$fidelity_complete_sum, 
            assess$final_school_id
            )

### Variance in fidelity by school
ggplot(
  assess[is.na(assess$final_school_id) == F, ],
  aes(x = final_school_id, 
      y = fidelity_complete_sum)
) +
  geom_boxplot(alpha = .7, fill = "gray") +
  theme_minimal()

# IntraClass Correlation
null_model <- lmer(
  fidelity_complete_sum ~ 1 +
  (1|final_school_id),
  data = assess
  
)
summary(null_model)
# ICC 0.3604234

### Variance in fedelity by school and condition
ggplot(
  assess[is.na(assess$final_school_id) == F, ],
  aes(x = final_school_id, 
      y = fidelity_complete_sum, 
      fill = condition_assignment)
) +
  geom_boxplot(alpha = .7) +
  theme_minimal()

# IntraClass Correlation
model_1 <- lmer(
  fidelity_complete_sum ~ 1 
  + condition_assignment +
  (1|final_school_id),
  data = assess
  
)
summary(model_1)
# ICC 0.364627 
# not a decrease, so school is a much bigger factor in fidelity than condition

### Variance in fedelity by school and implementation model (remote/in-person/) 
table(assess$movement)
# coding all 

ggplot(
  assess[is.na(assess$final_school_id) == F, ],
  aes(x = final_school_id, 
      y = fidelity_complete_sum, 
      fill = movement)
) +
  geom_boxplot(alpha = .7) +
  theme_minimal()

# Intraclass Correlation
model_2 <- lmer(
  fidelity_complete_sum ~ 1 
  + movement +
  (1|final_school_id),
  data = assess
  
)
summary(model_2)
# ICC 0.3923657
# not a decrease, so school is a much bigger factor in fidelity than implemntation type



```

### Teacher/Class
```{r}

# how many teachers/classes?
length(unique(assess$final_teacher_id))
length(unique(assess$final_teacher_class))
table(is.na(assess$final_teacher_class)) # 208

ggplot(
  assess[is.na(assess$final_school_id) == F, ],
  aes(x = final_teacher_class, 
      y = fidelity_complete_sum,
      fill = final_school_id,
      color = final_school_id
  )) +
  geom_boxplot(alpha = .7)

null_model2 <- lmer(
  fidelity_complete_sum ~ 1 
  + (1|final_school_id/final_teacher_class),
  data = assess
  
)
summary(null_model2)

# ICC 
# 0.288 teacher/school
# 0.3194 school



```

